---
title: "CompleteMePlease_StatisticalModellingSplicing_RiboMed2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

*Libraries*

```{r - Libraries}

library(psichomics)
library(cowplot)
library(ggplot2)
library(reshape2)
library(ggpubr)

library(RColorBrewer) #brewel.pal
library(scales)       #show_col only

```

*Variables*

Please find pre-processed files at: https://drive.google.com/drive/folders/17BhG4uMVVb8FC-m5an2zAzNUyfc2yKm0?usp=sharing
- GTEx_BrainMuscle_psis
- GTEx_BrainMuscle
- annot

**dataFolder** should be where downloaded .rds files are stored

```{r - Variables}

brain_color   <- "#1E90FF"
muscle_color  <- "#DC143C"
green         <- "#3CB371"
brain_pal     <- c(brewer.pal(n = 12, name = "Set3"), "white")
show_col(brain_pal)

beta_pal     <- brewer.pal(n = 5, name = "Set1")
show_col(beta_pal)

colLow  <- "#4169E1"
colHigh <- "#EF767A"

# Folder where pre-processed R files are stored locally
# :::::::::::::::::::::::::::::::::::::::::::::::::::::::
dataFolder <- "/Users/MarianaFerreira/Dropbox/iMM/Resources/"
# :::::::::::::::::::::::::::::::::::::::::::::::::::::::

GTEx_BrainMuscle_psis <- readRDS(file = paste0(dataFolder, "GTEx_BrainMuscle_psis.rds"))
GTEx_BrainMuscle      <- readRDS(file = paste0(dataFolder, "GTEx_BrainMuscle.rds"))
annot                 <- readRDS(file = paste0(dataFolder, "annot.rds"))

```

*Functions*

```{r - Functions}

findJunctionReads <- function (junctionQuant, annotation, minReads = 0){
  eventType = "SE"
  if (is.null(annotation$Gene)) 
    geneCol <- NULL
  else geneCol <- "Gene"
  coords <- rownames(junctionQuant)
  showStrand <- any(grepl("\\+|\\-", coords))
  annotation <- psichomics:::uniqueBy(annotation, "Chromosome", "Strand", 
                         "Constitutive exon 1 end", "Alternative exon 1 start", 
                         "Alternative exon 1 end", "Constitutive exon 2 start", 
                         geneCol)
  chr <- annotation$Chromosome
  strand <- annotation$Strand
  incAstr <- psichomics:::junctionString(chr, strand, annotation$`Constitutive exon 1 end`, 
                            annotation$`Alternative exon 1 start`, showStrand)
  incBstr <- psichomics:::junctionString(chr, strand, annotation$`Alternative exon 1 end`, 
                            annotation$`Constitutive exon 2 start`, showStrand)
  excAstr <- psichomics:::junctionString(chr, strand, annotation$`Constitutive exon 1 end`, 
                            annotation$`Constitutive exon 2 start`, showStrand)
  incA <- fastmatch::fmatch(incAstr, coords)
  incB <- fastmatch::fmatch(incBstr, coords)
  excA <- fastmatch::fmatch(excAstr, coords)
  excB <- 0
  nas <- is.na(incA) | is.na(incB) | is.na(excA)
  incA <- incA[!nas]
  incB <- incB[!nas]
  excA <- excA[!nas]
  if (length(incA) == 0) 
    return(NULL)
  multigene <- lapply(annotation$Gene, length) > 1
  gene <- annotation$Gene
  gene[multigene] <- lapply(gene[multigene], paste, collapse = "/")
  eventNames <- paste(sep = "_", eventType, chr, strand, 
                      annotation$`Constitutive exon 1 end`, annotation$`Alternative exon 1 start`, 
                      annotation$`Alternative exon 1 end`, annotation$`Constitutive exon 2 start`, 
                      gene)
  df <- cbind(incA, incB, excA)
  rownames(df) <- eventNames[!nas]
  return(df)
}

```

Learn about available *tissues* and *data types* for GTEx' splicing-related data

```{r [Do not run] Available tissues and datatypes}

# Download GTEx splicing data (version 7) from GoogleDrive: https://drive.google.com/drive/folders/1Z5Cibt8YgTnxfOfVCCq80KdHL-ChJ45o?usp=sharing

# psichomics::getGtexTissues()
# psichomics::getGtexTissues(folder = "/Users/MarianaFerreira/Dropbox/iMM/Resources/GTEx_splicing_v7")
# psichomics::getGtexDataTypes()

```

Load GTEx data for *Muscle* and *Brain* (v7)

```{r [Do not run] Load GTEx data for Muscle and Brain (v7)}

# GTEx_BrainMuscle <- psichomics::loadGtexData(data = c("sampleInfo", "subjectInfo", "junctionQuant"),
#                                              tissue = c("Brain", "Muscle"),
#                                              release = 7)

```

Load alternative splicing annotation (SE: skipped exon) and quantify PSIs

```{r [Do not run] Load alternative SE splicing annotation}

# psichomics::listSplicingAnnotations(species = "Human", assembly = "hg19", date = "2017")
# annot <- psichomics::loadAnnotation(annotation = listSplicingAnnotations(species = "Human", assembly = "hg19", date = "2017"))

# GTEx_BrainMuscle_psis <- quantifySplicing(annotation = annot,
#                                           junctionQuant = GTEx_BrainMuscle$GTEx$`Junction quantification`,
#                                           eventType = c("SE"),
#                                           minReads = 0)

```

Process GTEx PSIs
- remove events with at least one NA (keep ~30% of the SE table)

```{r Process GTEx PSIs}

# Identify rows (events) that have at least one NA
rowsAnyNA <- apply(GTEx_BrainMuscle_psis, 1, anyNA)

# Filter PSI table to keep only rows (events) with no NAs
psis <- GTEx_BrainMuscle_psis[!rowsAnyNA,]
cat(paste0("Fraction of events with no NAs: ", round(nrow(psis)/nrow(GTEx_BrainMuscle_psis), digits = 3)))

# match is required because Sample metadata contains more samples than those in PSI table
tissues       <- GTEx_BrainMuscle$GTEx_V7$`Sample metadata`$`Tissue Type (area of retrieval)`[match(colnames(psis),
                                                                                                    rownames(GTEx_BrainMuscle$GTEx_V7$`Sample metadata`))]
subtissues    <- GTEx_BrainMuscle$GTEx_V7$`Sample metadata`$`Tissue Type (detail)`[match(colnames(psis),
                                                                                         rownames(GTEx_BrainMuscle$GTEx_V7$`Sample metadata`))]

# Define the columns in PSI table that correspond to muscle and brain samples
muscle_cols   <- which(tissues == "Muscle")
brain_cols    <- which(tissues == "Brain")

```

**Ex1.** Check distribution of tissue areas for *Brain* samples

```{r Ex1. Distribution of brain sample sites}

# ::: Ex. 1 :::

```

**Ex2.** Inspect distribution of PSI values for *Brain* and *Muscle*

```{r Ex2. Inspect distribution of PSI values}

# ::: Ex. 2 :::

```

**Ex3.** Inspect the behaviour of PSI values across samples

- Are all events worth looking at?
- If not, how do we select the most relevant?

3.1. Calculate metrics that summarise information about PSIs across samples

```{r Ex3.1. Inspect behaviour of PSI values across samples}

# ::: Ex. 3.1 :::

# overallMedian <- apply(psis, 1, median)
# (...)

# Tip:
# GTEx_AllSE_BrainMuscle <- as.data.frame(cbind(::: ALL METRICS :::)

```

**Ex3.** Inspect the behaviour of PSI values across samples

- Are all events worth looking at?
- If not, how do we select the most relevant?

3.2. Plot and make justified decisions

```{r Ex3.2. Inspect behaviour of PSI values across samples (plot)}

# ::: Ex. 3.2 :::

```

**Ex4.** Filter to keep the most relevant ("alternative") events and check the resulting PSI distribution.

```{r Ex4. Filter to keep most relevant (alternative) events and plot}

# ::: Ex. 4 :::
# AllSE_BrainMuscle_Filtered <- GTEx_AllSE_BrainMuscle[::: FILTER CRITERIA :::,]
# filteredPsis <- :::: matching psi table :::

```

[Optional] Logit-transform PSIs

```{r [Optional] Logit-transform PSIs}

psistoPlot$logit <- car::logit(psistoPlot$value) 

ggplot(data = psistoPlot,
       aes(x = logit,
           fill = tissue)) +
  geom_density(color = "black",
               show.legend = FALSE) +
  scale_fill_manual(values = c(brain_color, muscle_color)) +
  xlab("logit(PSI)") +
  facet_wrap(. ~ tissue) +
  theme(text = element_text(size = 20))

```

Extract read count (inclusion and exclusion) data for considered events

NOTE: depends on tailor-made function *findJunctionReads* that finds inclusion and exclusion junctions (from GTEx Junction Quantification object) required for an event's PSI quantification

```{r Extract read count data}

readPositions <- findJunctionReads(annotation = annot$`Skipped exon`,
                                   junctionQuant = GTEx_BrainMuscle$GTEx_V7$`Junction quantification`,
                                   minReads = 0)

readsFromPsis <- as.data.frame(readPositions[match(rownames(filteredPsis), rownames(readPositions)),])

incA <- GTEx_BrainMuscle$GTEx_V7$`Junction quantification`[readsFromPsis$incA,]
incB <- GTEx_BrainMuscle$GTEx_V7$`Junction quantification`[readsFromPsis$incB,]
exc <- GTEx_BrainMuscle$GTEx_V7$`Junction quantification`[readsFromPsis$excA,]

inc <- (incA + incB)/2

inc <- inc[, match(colnames(filteredPsis), colnames(inc))]
exc <- exc[, match(colnames(filteredPsis), colnames(exc))]

cov <- inc + exc

```

Sum-up information about events under consideration
- Calculate metrics that summarise information about PSIs across samples, including for coverage

```{r Inspect behaviour of PSI and coverage values across samples}

psiTable <- filteredPsis
covTable <- cov

overallVariance       <- apply(psiTable, 1, var)
overallVarianceLogit  <- apply(car::logit(psiTable), 1, var)
overallMedian         <- apply(psiTable, 1, median)
overallMean           <- apply(psiTable, 1, mean)
maxPsi                <- apply(psiTable, 1, max)
minPsi                <- apply(psiTable, 1, min)

brainMedian           <- apply(psiTable[,brain_cols], 1, median)
brainMean             <- apply(psiTable[,brain_cols], 1, mean)
brainVariance         <- apply(psiTable[,brain_cols], 1, var)

muscleMedian          <- apply(psiTable[,muscle_cols], 1, median)
muscleMean            <- apply(psiTable[,muscle_cols], 1, mean)
muscleVariance        <- apply(psiTable[,muscle_cols], 1, var)

deltaPSI              <- brainMedian - muscleMedian
psiRange              <- maxPsi - minPsi

meanCoverageBrain     <- apply(covTable[,brain_cols], 1, mean)
meanCoverageMuscle    <- apply(covTable[,muscle_cols], 1, mean)
meanCoverageOverall   <- apply(covTable, 1, mean)

medianCoverageBrain   <- apply(covTable[,brain_cols], 1, median)
medianCoverageMuscle  <- apply(covTable[,muscle_cols], 1, median)
medianCoverageOverall <- apply(covTable, 1, median)

GTEx_FilteredSE_BrainMuscle <- as.data.frame(cbind(overallVariance, overallVarianceLogit, overallMedian, overallMean,
                                                   brainMedian, brainMean, brainVariance, medianCoverageBrain, meanCoverageBrain,
                                                   muscleMedian, muscleMean, muscleVariance, medianCoverageMuscle, meanCoverageMuscle,
                                                   medianCoverageOverall, meanCoverageOverall, deltaPSI, psiRange))

GTEx_FilteredSE_BrainMuscle$EventID <- rownames(GTEx_FilteredSE_BrainMuscle)

```

**Ex5.** Select *some* events of interest (OI). Can you predict their PSI distributions?

```{r Ex5. Select some events of interest}

eventsOI <- c("SE_1_+_168014465_168020130_168020171_168032859_DCAF6",
              "SE_16_+_2257346_2258211_2258335_2258451_MLST8",
              "SE_2_-_127819691_127818216_127818172_127816731_BIN1",
              "SE_17_-_1257505_1251302_1251256_1248793_YWHAE",
              "SE_1_+_11718421_11718590_11718685_11718793_FBXO44",
              "SE_1_-_37967405_37962337_37962308_37961519_MEAF6",
              "SE_19_+_10913034_10914315_10914344_10916592_DNM2",
              "SE_1_-_154142876_154141859_154141781_154130197_TPM3")

# ::: Ex. 5 :::

```

Check overall PSI distribution for selected events

```{r Check overall PSI distribution for selected events}

# function that plots PSIs from GTEx PSI table and plots jittered points and violin plots for all samples of a given eventID (optionally includes a Kruskal-Wallis test and a boxplot)
plotEventPsis <- function(eventID, includeKruskal, includeBoxplot){

  eventrow <- which(rownames(filteredPsis) == eventID)
  
  psistoPlot <- melt(filteredPsis[eventrow,])
  
  psistoPlot$tissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (area of retrieval)`[match(psistoPlot$variable,
                                                                                                       rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
  psistoPlot$subtissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (detail)`[match(psistoPlot$variable,
                                                                                               rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
  
  plot <- ggplot(data = psistoPlot,
         aes(x = tissue,
             y = value,
             group = tissue,
             color = tissue,
             fill = tissue)) +
    geom_violin(show.legend = FALSE,
                alpha = 0.5) +
    geom_jitter(alpha = 0.2,
                show.legend = FALSE) +
    scale_fill_manual(values = c(brain_color, muscle_color)) +
    scale_color_manual(values = c(brain_color, muscle_color)) +
    scale_y_continuous(breaks = seq(0, 1, 0.25), limits = c(0,1)) +
    xlab("tissue") +
    ylab("PSI") +
    labs(title = eventID) +
    theme(plot.title = element_text(size = 8), axis.title.x = element_blank())  +
    theme(text = element_text(size = 20))
  
  if(includeKruskal == TRUE){
    
    plot <- plot +
    stat_compare_means(method = "kruskal.test",
                       label.y = 1,
                       show.legend = FALSE)
    
  }else{}
  
  if(includeBoxplot == TRUE){
    
    plot <- plot +
    geom_boxplot(fill = NA,
                 outlier.shape = NA,
                 color = "black",
                 width = 0.25)
    
  }

  return(plot)
  
}

```

**Ex6.** What if we sample N (10) samples? And 3 samples?

```{r Ex6. What if we sample from these samples?}

# function that takes PSI table and generates jittered points and violin plots for a given eventID (optionally includes a Kruskal-Wallis test and a boxplot)
plotEventPsisAndSample <- function(eventID, sampleN, includeKruskal){

  eventrow <- which(rownames(filteredPsis) == eventID)
  
  psistoPlot <- melt(filteredPsis[eventrow,])
  
  psistoPlot$tissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (area of retrieval)`[match(psistoPlot$variable,
                                                                                                       rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
  psistoPlot$subtissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (detail)`[match(psistoPlot$variable,
                                                                                               rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
  
  sample <- psistoPlot[c(sample(x = which(psistoPlot$tissue == "Brain"), size = sampleN),
                         sample(x = which(psistoPlot$tissue == "Muscle"), size = sampleN)),]
  
  plot <- ggplot(data = psistoPlot,
         aes(x = tissue,
             y = value,
             group = tissue,
             color = tissue,
             fill = tissue)) +
    geom_violin(show.legend = FALSE,
                alpha = 0.5) +
    geom_jitter(alpha = 0.2,
                show.legend = FALSE) +
    geom_jitter(data = sample,
                color = "black",
                fill = "black",
                show.legend = FALSE) +
    scale_fill_manual(values = c(brain_color, muscle_color)) +
    scale_color_manual(values = c(brain_color, muscle_color)) +
    xlab("tissue") +
    ylab("PSI") +
    labs(title = eventID) +
    theme(plot.title = element_text(size = 8), axis.title.x = element_blank())
  
    if(includeKruskal == TRUE){
    
    plot <- plot +
      stat_compare_means(data = sample,
                       method = "kruskal.test",
                       label.y = 1,
                       show.legend = FALSE)
    }else{}
    
  return(plot)
  
}

# ::: Ex. 6 :::

```

**Ex7.** Inspect distribution of coverage (inc+exc) values for *Brain* and *Muscle*

```{r Ex7. Inspect distribution of coverage values}

# ::: Ex. 7 :::
  
```

**Ex8.** Simulate PSI-associated confidence intervals from a proportion's test.

```{r Ex8. Simulate PSI-associated confidence intervals from a proportion's test}

# ::: Ex. 8 :::

```

**Ex9.** Explore R stats beta distribution functions

```{r Ex9. Explore R stats beta distribution functions}

# df <- as.data.frame(cbind(c(points1, # ::: (...) :::),
#                           c(dens1, # ::: (...) :::)),
#                           c(rep("label1", times = 1000), # ::: (...) :::))))

# ::: Ex. 9 :::

```

**Ex10.** Simulate beta distributions with different coverage levels

```{r Ex10. Simulate beta distributions with different coverage levels}

# ::: Ex. 10 :::

```

Use beta-distribution approach to model inclusion levels for these events

```{r Use beta-distribution approach to model inclusion levels from sampled replicates}

N <- 3
names1 <- paste0("Brain", 1:N)
names2 <- paste0("Muscle", 1:N)

# Column in reference table
i <- which(rownames(filteredPsis) == eventsOI[2])

# :::::::::::
# Get original PSIs and median PSIs
# :::::::::::

psisBrain  <- as.numeric(as.vector(filteredPsis[i,brain_cols]))
psisMuscle <- as.numeric(as.vector(filteredPsis[i,muscle_cols]))

psiRefBrain  <- GTEx_FilteredSE_BrainMuscle$brainMedian[i]
psiRefMuscle <- GTEx_FilteredSE_BrainMuscle$muscleMedian[i]


originalpsi <- ggplot() +
  geom_density(aes(x = psisBrain),
               color = NA,
               alpha = 0.5,
               fill = brain_color) +
  geom_rug(aes(x = psisBrain),
           alpha = 0.5,
           color = brain_color) +
  geom_vline(xintercept = psiRefBrain,
             color = brain_color,
             size = 1) +
  geom_density(aes(x = psisMuscle),
               color = NA,
               alpha = 0.5,
               fill = muscle_color) +
  geom_rug(aes(x = psisMuscle),
           alpha = 0.5,
           color = muscle_color) +
  geom_vline(xintercept = psiRefMuscle,
             color = muscle_color,
             size = 1) +
  xlab("PSI") +
  scale_x_continuous(limits = c(0,1),
                     breaks = seq(0,1,0.25)) +
  theme(axis.title = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 12)) +
  labs(title = "Original PSIs")

# :::::::::::
# Get original coverage and mean overall coverage
# :::::::::::

covsBrain  <- as.numeric(as.vector(cov[i,brain_cols]))
covsMuscle <- as.numeric(as.vector(cov[i,muscle_cols]))

covRefBrain <- GTEx_FilteredSE_BrainMuscle$medianCoverageBrain[i]
covRefMuscle <- GTEx_FilteredSE_BrainMuscle$medianCoverageMuscle[i]

originalcov <- ggplot() +
  geom_histogram(aes(x = covsBrain),
                 alpha = 0.5,
                 fill = brain_color,
                 color = NA,
                 bins = 100) +
  geom_rug(aes(x = covsBrain),
           alpha = 0.5,
           color = brain_color) +
  geom_vline(xintercept = covRefBrain,
             color = brain_color,
             size = 1) +
  geom_histogram(aes(x = covsMuscle),
                 alpha = 0.5,
                 fill = muscle_color,
                 color = NA,
                 bins = 100) +
  geom_rug(aes(x = covsMuscle),
           alpha = 0.5,
           color = muscle_color) +
  geom_vline(xintercept = covRefMuscle,
             color = muscle_color,
             size = 1) +
  scale_x_log10(limits = c(1,10000),
                breaks = c(1,10,100,1000)) +
  xlab("inc+exc") +
  theme(axis.title = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 12)) +
  labs(title = "Original coverage")

  # Sample N samples (use for both PSI and coverage)
  
  allPsis <- filteredPsis[i,]
  allCovs <- cov[i,]

  psisToSampleFrom <- melt(allPsis)
  psisToSampleFrom$tissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (area of retrieval)`[match(psisToSampleFrom$variable,
                                                                                                       rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
  psisToSampleFrom$subtissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (detail)`[match(psisToSampleFrom$variable,
                                                                                               rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
  
  brain_samples <- psisToSampleFrom$variable[sample(x = which(psisToSampleFrom$tissue == "Brain"), size = N)]
  muscle_samples <- psisToSampleFrom$variable[sample(x = which(psisToSampleFrom$tissue == "Muscle"), size = N)]
  
  brain_covs <- as.numeric(as.vector(cov[i,which(colnames(cov) %in% brain_samples)]))
  brain_inc <- as.numeric(as.vector(inc[i,which(colnames(inc) %in% brain_samples)]))
  brain_exc <- as.numeric(as.vector(exc[i,which(colnames(exc) %in% brain_samples)]))
  
  muscle_covs <- as.numeric(as.vector(cov[i,which(colnames(cov) %in% muscle_samples)]))
  muscle_inc <- as.numeric(as.vector(inc[i,which(colnames(inc) %in% muscle_samples)]))
  muscle_exc <- as.numeric(as.vector(exc[i,which(colnames(exc) %in% muscle_samples)]))
  
  artif1 <- c()
  artif2 <- c()
  
  artif_sample1 <- c()
  artif_sample2 <- c()
  
  for(j in 1:N){
    
    artif_sample1   <- c(artif_sample1, rep(names1[j], times = 500))
    artif_sample2   <- c(artif_sample2, rep(names2[j], times = 500))
    
    artif1          <- c(artif1,
                         rbeta(500, shape1 = brain_inc[j], shape2 = brain_exc[j]))
    
    artif2          <- c(artif2,
                         rbeta(500, shape1 = muscle_inc[j], shape2 = muscle_exc[j]))

  }
  
  median1 <- median(artif1)
  median2 <- median(artif2)
  
  names(artif1) <- artif_sample1
  names(artif2) <- artif_sample2
  
  p2 <- ggplot() +
    geom_density(aes(x = as.numeric(as.vector(artif1)),
                     group = names(artif1)),
                 alpha = 0.3,
                 color = brain_color,
                 fill = brain_color) +
    geom_density(aes(x = as.numeric(as.vector(artif2)),
                     group = names(artif2)),
                 alpha = 0.3,
                 color = muscle_color,
                 fill = muscle_color) +
    xlab("PSI") +
    scale_x_continuous(limits = c(0,1),
                       breaks = seq(0,1,0.25)) +
    geom_vline(xintercept = median1,
               color = brain_color,
               size = 1) +
    geom_vline(xintercept = median2,
               color = muscle_color,
               size = 1) +
    theme(legend.position = c(0.8,0.8),
          legend.text = element_text(size = 20),
          legend.key = element_blank(),
          legend.background = element_blank(),
          axis.title = element_text(size = 14),
          axis.text.y = element_text(size = 14),
          axis.text.x = element_text(size = 12)) +
    labs(title = "Beta dist. PSI")
    
  multiplot <- plot_grid(originalpsi, originalcov, p2, nrow = 3, align = "v", rel_heights = c(1,1,3))
  multiplot

```

Use beta-distribution approach to model inclusion levels for these events (function mode)

```{r Use beta-distribution approach to model inclusion levels from sampled replicates}

# Select fixed N replicates
nreps <- 5
brain_nrep_cols <- sample(x = which(tissues == "Brain"), size = nreps)
brain_samples   <- colnames(filteredPsis)[brain_nrep_cols]

muscle_nrep_cols  <- sample(x = which(tissues == "Muscle"), size = nreps)
muscle_samples    <- colnames(filteredPsis)[muscle_nrep_cols]

# ::::::::::::::::

gtexNrepsBetas <- function(row){
  
  nrepBetas <- list()

brain_covs  <- as.numeric(as.vector(cov[row, brain_samples]))
brain_inc   <- as.numeric(as.vector(inc[row, brain_samples]))
brain_exc   <- as.numeric(as.vector(exc[row, brain_samples]))

muscle_covs <- as.numeric(as.vector(cov[row, muscle_samples]))
muscle_inc  <- as.numeric(as.vector(inc[row, muscle_samples]))
muscle_exc  <- as.numeric(as.vector(exc[row, muscle_samples]))

artif_brain <- c()
artif_muscle <- c()

artif_sample_brain  <- c()
artif_sample_muscle <- c()

for(j in 1:nreps){
  
  artif_sample_brain  <- c(artif_sample_brain, rep(brain_samples[j], times = 500))
  artif_sample_muscle <- c(artif_sample_muscle, rep(muscle_samples[j], times = 500))
  
  artif_brain         <- c(artif_brain,
                           rbeta(500, shape1 = brain_inc[j], shape2 = brain_exc[j]))
  
  artif_muscle        <- c(artif_muscle,
                           rbeta(500, shape1 = muscle_inc[j], shape2 = muscle_exc[j]))
  
}

median1 <- median(artif_brain)
median2 <- median(artif_muscle)

names(artif_brain)  <- artif_sample_brain
names(artif_muscle) <- artif_sample_muscle

prob <- c()
seq  <- seq(-1, 1, length.out = 1001)

artifA <- sample(x = artif_brain, size = length(artif_brain))
artifB <- sample(x = artif_muscle, size = length(artif_muscle))

p_zero_Fast <- abs(length(which(artifA - artifB > 0))/length(artifA)-0.5)+0.5

nrepBetas[[1]] <- median(artif_brain)
nrepBetas[[2]] <- median(artif_muscle)
nrepBetas[[3]] <- p_zero_Fast

nrepBetas[[4]] <- artif_brain
nrepBetas[[5]] <- artif_muscle

names(nrepBetas) <- c("BrainMedianPsi", "MuscleMedianPsi", "Pdiff", "BrainPoints", "MusclePoints")
return(nrepBetas)

  
}

  gtexBetas <- lapply(1:nrow(filteredPsis), function(x) gtexNrepsBetas(row = x))
  names(gtexBetas) <- rownames(filteredPsis)
  
diffTable <- AllSE_BrainMuscle_Filtered
diffTable$BrainMedianPsi      <- as.numeric(as.vector(lapply(1:length(gtexBetas), function(x) gtexBetas[[x]][[1]])))
diffTable$MuscleMedianPsi     <- as.numeric(as.vector(lapply(1:length(gtexBetas), function(x) gtexBetas[[x]][[2]])))
diffTable$deltaBrainToMuscle  <- diffTable$BrainMedianPsi - diffTable$MuscleMedianPsi
diffTable$Pdiff               <- as.numeric(as.vector(lapply(1:length(gtexBetas), function(x) gtexBetas[[x]][[3]])))

ggplot(data = diffTable,
       aes(x = deltaBrainToMuscle,
           y = Pdiff,
           color = deltaPSI)) +
  geom_point() +
  scale_color_gradient2(name = expression(Delta*PSI),
                        low = muscle_color,
                        mid = "gray",
                        high = brain_color) +
  geom_point(data = diffTable[eventsOI,],
             color = "black",
             fill = NA,
             shape = 21) +
  # xlab(expression(abs(Delta*PSI[Brain-Muscle]))) +
  xlab(expression(PSI[Brain]-PSI[Muscle])) +
  ylab("Probability of differential splicing") +
  labs(caption = paste0(nrow(filteredPsis), " exon skipping (SE) events")) +
  theme(text = element_text(size = 20))

```

Inspect events of interest

```{r}

# Study the "outliers" of the volcano
ggplot(data = diffTable,
       aes(x = deltaBrainToMuscle,
           y = Pdiff,
           color = deltaPSI)) +
  geom_point() +
  geom_abline(intercept = c(0.5, 0.5),
              slope = c(-1, 1)) +
  scale_color_gradient2(name = expression(Delta*PSI),
                        low = muscle_color,
                        mid = "gray",
                        high = brain_color) +
  geom_point(data = examples,
             color = "black",
             fill = NA,
             shape = 21) +
  # xlab(expression(abs(Delta*PSI[Brain-Muscle]))) +
  xlab(expression(PSI[Brain]-PSI[Muscle])) +
  ylab("Probability of differential splicing") +
  labs(caption = paste0(nrow(filteredPsis), " exon skipping (SE) events"))


examples <- rownames(diffTable[which((diffTable$Pdiff<diffTable$deltaBrainToMuscle+0.5 | diffTable$Pdiff < -diffTable$deltaBrainToMuscle+0.5) &
                              diffTable$Pdiff > 0.7 & diffTable$Pdiff < 1),])

extreme_examples <- rownames(diffTable[which((diffTable$Pdiff < -diffTable$deltaBrainToMuscle+0.5) &
                                               diffTable$Pdiff > 0.7 & diffTable$Pdiff < 0.9),])

```

Plot specific cases

```{r}

# Select a random event to plot from the subset of "events of interest" and "examples"
i <- sample(which(rownames(filteredPsis) %in% c(eventsOI, extreme_examples)), size = 1)
# ::::::::::::::::::::::::::

volcanoInd <- ggplot(data = diffTable,
                     aes(x = deltaBrainToMuscle,
                         y = Pdiff,
                         color = deltaPSI)) +
  geom_point(show.legend = FALSE,
             alpha = 0.5) +
  scale_color_gradient2(name = expression(Delta*PSI),
                        low = muscle_color,
                        mid = "gray",
                        high = brain_color) +
  geom_point(data = diffTable[i,],
             color = "black") +
  xlab(expression(PSI[Brain]-PSI[Muscle])) +
  ylab("Probability of differential splicing") +
  labs(title = rownames(diffTable)[i])

psistoPlot <- melt(filteredPsis[i,])

psistoPlot$tissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (area of retrieval)`[match(psistoPlot$variable,
                                                                                                     rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]
psistoPlot$subtissue <- GTEx_BrainMuscle$GTEx$`Sample metadata`$`Tissue Type (detail)`[match(psistoPlot$variable,
                                                                                             rownames(GTEx_BrainMuscle$GTEx$`Sample metadata`))]

usedSamples <- c(unique(names(gtexBetas[[i]][[4]])), unique(names(gtexBetas[[i]][[5]])))
sample      <- psistoPlot[match(usedSamples, psistoPlot$variable),]

plotPsis <- ggplot(data = psistoPlot,
                   aes(x = tissue,
                       y = value,
                       group = tissue,
                       color = tissue,
                       fill = tissue)) +
  geom_violin(show.legend = FALSE,
              alpha = 0.5) +
  geom_jitter(alpha = 0.2,
              show.legend = FALSE) +
  geom_jitter(data = sample,
              color = "black",
              fill = "black",
              show.legend = FALSE) +
  scale_fill_manual(values = c(brain_color, muscle_color)) +
  scale_color_manual(values = c(brain_color, muscle_color)) +
  xlab("tissue") +
  ylab("PSI") +
  theme(plot.title = element_text(size = 8),
        axis.title.x = element_blank()) +
  stat_compare_means(data = sample,
                     method = "kruskal.test",
                     label.y = 1,
                     show.legend = FALSE)

artif1 <- gtexBetas[[i]][[4]]
artif2 <- gtexBetas[[i]][[5]]
median1 <- gtexBetas[[i]][[1]]
median2 <- gtexBetas[[i]][[2]]

pbrain <- ggplot() +
  geom_jitter(aes(x = factor(names(artif1)),
                  y = as.numeric(as.vector(artif1))),
              alpha = 0.2,
              color = brain_color,
              fill = brain_color) +
  geom_hline(yintercept = median1,
             color = brain_color) +
  xlab("") +
  scale_y_continuous(limits = c(0,1),
                     breaks = seq(0,1,0.25)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x.bottom = element_blank(),
        axis.title.y = element_blank())

pmuscle <- ggplot() +
  geom_jitter(aes(x = factor(names(artif2)),
                  y = as.numeric(as.vector(artif2))),
              alpha = 0.2,
              color = muscle_color,
              fill = muscle_color) +
  geom_hline(yintercept = median2,
             color = muscle_color) +
  xlab("") +
  scale_y_continuous(limits = c(0,1),
                     breaks = seq(0,1,0.25)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x.bottom = element_blank(),
        axis.title.y = element_blank())

multiplotRow  <- plot_grid(plotPsis, pbrain, pmuscle, ncol = 3, align = "h")
multiplot     <- plot_grid(volcanoInd, multiplotRow, nrow = 2)

multiplot
rownames(diffTable)[i]
rownames(diffTable)[i] %in% eventsOI

```

